
---
title: Test for cell fractions using the compositional data analysis (T cell)
output: html_notebook
---



Functions:

```{r}
simple_proportion_test <- function(data,  # rowns - samples; columns - cell types
                              samplegroups,  # list of target and control groups
                              cell_list1,  # array of cell types of interest
                              cell_list2=NULL,  # Alternative set of cells; NULL when "others"
                              quantile = 0.95,  # Quantile for power
                              niter = 1000,  # Number of bootstrap iterations
                              n_cell_counts = 500,  # Number of cells to remain
                              alternative = 'two.sided',
                              nseed = 239,
                              n.cores=1) {

  set.seed(nseed)
  rl <- unlist(pagoda2:::papply(1:niter,function(i) {
    # bootstrap resampling of the cells in each sample, producing resampled version of the data table
    data <- t(apply(data,1,function(x) table(factor(sample(rep(colnames(data),x),sum(x),replace = T),levels=colnames(data)))))
    
    # run the test
    if(is.null(cell_list2)) {
      df <- cbind(rowSums(data[,cell_list1,drop=F]),rowSums(data[,setdiff(colnames(data),cell_list1),drop=F])); 
      colnames(df) <- c(paste(cell_list1,collapse = '+'),'other')
    } else {
      df <- cbind(rowSums(data[,cell_list1,drop=F]),rowSums(data[,cell_list2,drop=F])); 
      colnames(df) <- c(paste(cell_list1,collapse = '+'),paste(cell_list2,collapse = '+'))
    }
    
    ratio <- df[,1]/rowSums(df); 
    ratio[is.nan(ratio)] <- NA;
    wilcox.test(ratio[samplegroups$trgt], ratio[samplegroups$ctrl],alternative = alternative)$p.value
  },n.cores=n.cores,mc.preschedule=T))
  
  return(list(pvals=rl,q_pval=max(quantile(rl,1-quantile),quantile(rl,quantile))))
}
```




```{r}

library(reshape2)
load('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision/F3/F3_ano.RData')

#  anoCell.T cell types
#  Tcell.col  color palette for cell types
#  fraction.T
#  allsample.T
fraction.palette=readRDS('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision/data/fraction.palette.rds')
fraction.palette2=c('grey',fraction.palette)
names(fraction.palette2)[1]='Healthy'


anoCell2=anoCell.T
anoCell2[anoCell2 %in% c('CTL-1','CTL-2')]='CTL'

ano2=data.frame('Group'=anoCell2,'sample'=allsample.T[names(anoCell2)])

# Annotation vs sample
tmp2 <- acast(ano2, Group ~ sample, fun.aggregate=length)
#head(tmp2)
# Normalise for the number of cells in each library
tmp3 <- (sweep(tmp2, 2, colSums(tmp2), FUN='/'))
tmp4 <- melt(tmp3)
#head(tmp4)
names(tmp4) <- c('annot', 'sample','pc.of.sample')
#head(tmp4)

tmp4$Fraction=apply(tmp4,1,function(x) strsplit(x[2],'-')[[1]][2])
tmp4$Fraction=ordered(as.factor(tmp4$Fraction), levels = c( "Healthy", "Benign", "Distal","Involved","Tumor"))

```



```{r fig.width=7, fig.height=4}
require(ggpubr)
### barplot
lis=list()
for(i in unique(anoCell2) ){
  tmp=tmp4[tmp4[,1]==i,]
  sig=compare_means(pc.of.sample ~ Fraction,  data = tmp)
  sig$cell=i
  sig=sig[sig$p.signif!='ns',]
  print(i)
  print(sig)
  p1 <- ggplot(tmp, aes(x=Fraction,fill=Fraction,y=pc.of.sample)) + geom_boxplot(outlier.shape = -1,width=0.3,position=position_dodge(width=0.1))+ facet_wrap(~annot,scales="free_y")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(axis.text.x=element_blank()) + xlab("")  + ylab("")
  #p1=p1+ geom_point(data = tmp, size = 0.6, shape = 21, position = position_jitterdodge()) 
  p1=p1+theme(strip.background = element_rect(fill=Tcell.col[i]),strip.text = element_text(colour = 'black'))+ theme(legend.position="none")
  p1=p1+stat_summary(fun.y=median,linetype="dotted", geom="smooth",size=0.8, aes(group=0),color='gray30')+scale_fill_manual(values=fraction.palette2)
  lis[[i]]=p1
  #p1
  i=sub('/','_',i)
  ggsave(paste(i,'.pdf',sep=''),p1,w=2.3,h=2.2)

}


b=  cowplot::plot_grid(plotlist=lis, ncol=4, nrow=2)

b


```





The First step:
Preparing the data: count table and 4 groups of samples (Involved, Noninvolved, Tumor, Whole)


```{r}
data=table(allsample.T[names(anoCell2)],anoCell2)

head(data)
s_names = rownames(data)

samplegroups <- list(
  inv = s_names[grepl("Involved", s_names, fixed = TRUE)],
  noninv = s_names[grepl("Distal", s_names, fixed = TRUE)],
  tum = s_names[grepl("Tumor", s_names, fixed = TRUE)],
  whl = s_names[grepl("Benign", s_names, fixed = TRUE)],
  heal =s_names[grepl("Healthy", s_names, fixed = TRUE)]
)

samplegroups
```











Example 1:

Compare "one cell type" vs "others" between two groups of samples 

Compare CTL vs others between tumor and Distal


Using simple bootstrap resampling
```{r}


samplegroups$ctrl = samplegroups$noninv  # Distal

samplegroups$trgt = samplegroups$tum  # Tumor
#samplegroups

cell_list1 = c('CTL')
quantile = 0.99
alternative = 'greater'

print(cell_list1)
print(quantile)
print(samplegroups$ctrl)
print(samplegroups$trgt)

res = simple_proportion_test(data,
                    samplegroups,
                    cell_list1,  # array of cell types of interest
                    quantile = quantile,  # Quantile for power
                    alternative = alternative,
                    n.cores=10)

res$q_pval
hist(log(res$pvals))
```




Using simple bootstrap resampling
Compare NK cells vs others between Distal and Benign

```{r}


samplegroups$ctrl = samplegroups$whl  

samplegroups$trgt = samplegroups$noninv  
#samplegroups

cell_list1 = c('NK')
quantile = 0.99
alternative = 'greater'

print(cell_list1)
print(quantile)
print(samplegroups$ctrl)
print(samplegroups$trgt)

res = simple_proportion_test(data,
                    samplegroups,
                    cell_list1,  # array of cell types of interest
                    quantile = quantile,  # Quantile for power
                    alternative = alternative,
                    n.cores=10)

res$q_pval
hist(log(res$pvals))
```







Using simple bootstrap resampling

Compare CD4+ Naive vs others between Healthy and Tumor fraction

```{r}


samplegroups$ctrl = samplegroups$tum  

samplegroups$trgt = samplegroups$heal  
#samplegroups

cell_list1 = c("CD4+ Naive")
quantile = 0.99
alternative = 'greater'

print(cell_list1)
print(quantile)
print(samplegroups$ctrl)
print(samplegroups$trgt)

res = simple_proportion_test(data,
                    samplegroups,
                    cell_list1,  # array of cell types of interest
                    quantile = quantile,  # Quantile for power
                    alternative = alternative,
                    n.cores=10)

res$q_pval
hist(log(res$pvals))
```













Using simple bootstrap resampling

Compare Th1/17 Naive vs others between Tumor and Distal fraction

```{r}


samplegroups$ctrl = samplegroups$noninv  # Distal

samplegroups$trgt = samplegroups$tum  # Tumor
#samplegroups

cell_list1 = c("Th1/17")
quantile = 0.99
alternative = 'greater'

print(cell_list1)
print(quantile)
print(samplegroups$ctrl)
print(samplegroups$trgt)

res = simple_proportion_test(data,
                    samplegroups,
                    cell_list1,  # array of cell types of interest
                    quantile = quantile,  # Quantile for power
                    alternative = alternative,
                    n.cores=20)

res$q_pval
hist(log(res$pvals))
```










