---
title: "R Avearge signature score "
output: html_notebook
---

# load data 
```{r}
source('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision/Lib.r')
source('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision/utils.R')

#load('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision/F2/F2.sample.RData')
library(ggpubr)


```

#  caculate avearge expression of signature gene set
```{r}

#' caculate avearge expression of signature gene set
#' @param anoCell factor of cell types
#' @param m2 signature gene set
#' @param aexp expression matrix 
#' @param fraction factor of fractions
#' @param anoSample vector of samples 
#' @param Bootstraps.quantile.cell  sampling cell  
#' @param Bootstraps.quantile.gs  sampling gene set  

Signature_score=function(anoCell,m2,aexp,fraction,anoSample,min.num.cell=10,magnitude.normal = NULL,Bootstraps.quantile.cell=NULL,Bootstraps.quantile.gs=NULL){
  m2=intersect(m2,colnames(aexp))
  cname=names(anoCell)
  if (!is.null(Bootstraps.quantile.cell)){
    Bootstraps.quantile.cell=ceiling(length(cname)*Bootstraps.quantile.cell)
    print(Bootstraps.quantile.cell)
    set.seed(sample(10000,1))
    cname=sample(cname,Bootstraps.quantile.cell,replace=T)
  }
  
  if (!is.null(Bootstraps.quantile.gs)){
    set.seed(sample(10000,1))
    Bootstraps.quantile.gs=ceiling(length(m2)*Bootstraps.quantile.gs)
    m2=sample(m2,Bootstraps.quantile.gs,replace=T)
    set.seed(NULL)
    
  }
  
  
  exp=as.matrix(aexp[cname,m2])
  print(dim(exp))
  mmax=apply(exp,2,function(x) max(x))
  exp=exp[,mmax>0]
  
  if (!is.null(magnitude.normal)) {
    exp = apply(exp, 2, function(x) x/max(x))
  }  
  m2score=rowMeans(as.matrix(exp))
  summary(m2score)
  
  m2score_scale=scale(m2score)
  names(m2score_scale)=names(m2score)
  print(summary(m2score))
  
  cname=names(m2score)
  
  
  ttype=paste(anoCell[cname],'|',anoSample[cname],'|',fraction[cname],sep='')
  names(ttype)=cname
  
  if (!is.null(min.num.cell)){
    sel= table(ttype) %>% .[.>min.num.cell] %>% names()
    ttype=ttype[ttype %in% sel]
  }
  # print(table(ttype))
  
  m2score=m2score[names(ttype)]
  score.sample=tapply(m2score,ttype,mean)
  
  ano=lapply(sn(names(score.sample)),function(x) strsplit(x,'[|]')[[1]])
  cell=unlist(lapply(ano,function(x) x[1]))
  anoSample=unlist(lapply(ano,function(x) x[2]))
  faction=unlist(lapply(ano,function(x) x[3]))
  
  dat2=data.frame('score'=score.sample,'cell'=cell,'fraction'=faction,'sample'=anoSample)
  dat2$name=paste(as.character(dat2[,'cell']),as.character(dat2[,'sample']))
  dat2$fraction=ordered(as.factor(dat2$fraction), levels = levels(fraction))   
  dat2$cell=ordered(as.factor(dat2$cell), levels = levels(anoCell))   
  
  #dat2=dat2[dat2$fraction=='Tumor',]
  return(dat2)
}



SignatureScoreBootstraps<-function(anoCell,m2,aexp,fraction,anoSample,min.num.cell=10,magnitude.normal = NULL,n.bootstraps=0,Bootstraps.quantile.cell=NULL,Bootstraps.quantile.gs=NULL,Bootstraps.cutoff=0.99){
  #nname=names(fraction[fraction=='Tumor'])
  #m2=glist$TIM
  #Bootstraps.quantile.gs=NULL
  #Bootstraps.quantile.cell=0.9
  #n.bootstraps=2000
  #Bootstraps.cutoff=0.99
  
  df=Signature_score(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE)
  fsig=compare_means(score ~ cell,  data = df)
  fsig=as.data.frame(fsig)
  rownames(fsig)=paste(fsig$group1,fsig$group2)
  
  fscore <- mclapply(sn(seq(n.bootstraps)), function(p) {
    df=Signature_score(anoCell[nname],m2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,
                       Bootstraps.quantile.cell=Bootstraps.quantile.cell,Bootstraps.quantile.gs=Bootstraps.quantile.gs)
    sig=compare_means(score ~ cell,  data = df)
    sig=as.data.frame(sig)
    rownames(sig)=paste(sig$group1,sig$group2)
    list('sig'=sig)
  },mc.cores=32)
  
  
  Bootstraps.pvalue=unlist(lapply(sn(rownames(fsig)),function(x) {
    t.score=unlist(lapply(fscore,function(d) d$sig[x,'p']))
    quantile(t.score,Bootstraps.cutoff)
    }))
  
  
  fsig$Bootstraps.pvalue=Bootstraps.pvalue
  
  return(list('df'=df,'fsig'=fsig))
}



```



# test on TIM signature socre (tumor fraction)
 Inflammatory signature are significant high in TIM population 
```{r fig.width=5, fig.height=5}
nname=names(fraction[fraction=='Tumor'])
table(anoCell[nname])

df=Signature_score(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE)

ylab='Inflammatory signature'
fsig=compare_means(score ~ cell,  data = df)
fsig



```


plot
```{r fig.width=5, fig.height=5}
p=drawBoxplot(df,ylab,myeloid.col)   # fraction.palette

```




# Bootstraps for cell size 
3000 times  
```{r}

sel=c('Mono-1 Mono-2','Mono-1 Mono-3','Mono-2 Mono-3','Mono-2 TIM','Mono-3 TIM','Mono-1 TIM')

df2=SignatureScoreBootstraps(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=3000,Bootstraps.quantile.cell=1,Bootstraps.quantile.gs=NULL,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]

```


# Bootstraps for cell size 
5000 times 

```{r fig.width=5, fig.height=5}


df2=SignatureScoreBootstraps(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=3000,Bootstraps.quantile.cell=1,Bootstraps.quantile.gs=NULL,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]

```




# Bootstraps for gene num 
3000 times

```{r}

df2=SignatureScoreBootstraps(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=3000,Bootstraps.quantile.cell=NULL,Bootstraps.quantile.gs=1,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]

```









#  diagnose figure 
```{r}
  nname=names(fraction[fraction=='Tumor'])
  m2=glist$TIM
  Bootstraps.quantile.gs=NULL
  Bootstraps.quantile.cell=1
  n.bootstraps=2000
  Bootstraps.cutoff=0.99
  Bootstraps.quantile.gs=NULL
  
  df=Signature_score(anoCell[nname],glist$TIM,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE)
  fsig=compare_means(score ~ cell,  data = df)
  fsig=as.data.frame(fsig)
  rownames(fsig)=paste(fsig$group1,fsig$group2)
  
  fscore <- mclapply(sn(seq(n.bootstraps)), function(p) {
    df=Signature_score(anoCell[nname],m2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,
                       Bootstraps.quantile.cell=Bootstraps.quantile.cell,Bootstraps.quantile.gs=Bootstraps.quantile.gs)
    sig=compare_means(score ~ cell,  data = df)
    sig=as.data.frame(sig)
    rownames(sig)=paste(sig$group1,sig$group2)
    list('sig'=sig)
  },mc.cores=32)
  
  
  
```

```{r}

lis=list()
lis[['Mono1 vs Mono2']]=-log10(unlist(lapply(fscore,function(x) x$sig[1,'p'])))
lis[['Mono1 vs Mono3']]=-log10(unlist(lapply(fscore,function(x) x$sig[2,'p'])))
lis[['Mono2 vs TIM']]=-log10(unlist(lapply(fscore,function(x) x$sig[8,'p'])))
lis[['Mono3 vs TIM']]=-log10(unlist(lapply(fscore,function(x) x$sig[11,'p'])))




mat=do.call(cbind,lis)
colnames(mat)=names(lis)
dat=melt(mat)


p1 <- ggplot(dat, aes(x=Var2,fill=Var2,y=value)) + geom_boxplot(outlier.shape = -1,width=0.5,position=position_dodge(width=0.3)) +theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("")  + ylab('-log 10 Pvalue')+ theme(legend.position="none")
p1=p1+ geom_point(data = dat,color=adjustcolor(1,alpha=0.3),size=0.1, position = position_jitterdodge(0.3)) 
p1

```

# test on M2 signature socre (tumor fraction)

```{r}
nname=names(fraction[fraction=='Tumor'])
table(anoCell[nname])

df=Signature_score(anoCell[nname],glist$M2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE)

ylab='Inflammatory signature'
fsig=compare_means(score ~ cell,  data = df)
fsig
```


plot
```{r fig.width=5, fig.height=5}
p=drawBoxplot(df,ylab,myeloid.col)   # fraction.palette

```








# Bootstraps for cell size 
3000 times
```{r}

sel=c('Mono-1 Mono-2','Mono-1 Mono-3','Mono-2 Mono-3','Mono-2 TAM','Mono-3 TAM','Mono-1 TAM')

df2=SignatureScoreBootstraps(anoCell[nname],glist$M2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=3000,Bootstraps.quantile.cell=1,Bootstraps.quantile.gs=NULL,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]

```

5000 times

```{r}
sel=c('Mono-1 Mono-2','Mono-1 Mono-3','Mono-2 Mono-3','Mono-2 TAM','Mono-3 TAM','Mono-1 TAM')

df2=SignatureScoreBootstraps(anoCell[nname],glist$M2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=5000,Bootstraps.quantile.cell=1,Bootstraps.quantile.gs=NULL,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]
```



# Bootstraps for  gene num 

3000 times 

```{r}

df2=SignatureScoreBootstraps(anoCell[nname],glist$M2,p2$counts,fraction,samplef,min.num.cell=5,magnitude.normal = TRUE,n.bootstraps=3000,Bootstraps.quantile.cell=NULL,Bootstraps.quantile.gs=1,Bootstraps.cutoff=0.99)
df2$fsig[sel,c('group1','group2','p','Bootstraps.pvalue')]

```








