---
title: "R Notebook S1"
output: html_notebook
---


# load data and show embedding 
```{r}

library(Seurat)

library(ggplot2)
library(cowplot)
library(patchwork)

setwd('/home/meisl/Workplace/BMME/Revision/R1')

source('/d0-mendel/home/meisl/Workplace/BMME/Figures/revision//Lib.r')

#load('R1.RData')
```



```{r}
# load('R1.RData)

# load conos data and cell annotation 

# pagoda object for Tumor fraction; basicP2proc() can generatepagoda object
p2T=readRDS('../TumorFraction.p2.0406.rds')

scon=readRDS('../data/BMET_conos.all.rds')

# cell annotation; 
# typefc2;fractionf;samplef
load('../data/ano.RData')

typefc2=as.factor(typefc2)
cellano=typefc2[grepl('Tumor',names(typefc2))]

# load color color palette for cell type 

cellcol=readRDS('../data/cellType.color.rds')


pf <- function(n) { cellcol[1:n]}

```


```{r fig.width=5, fig.height=5}
p0 <- scon$plotGraph(raster=T,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.8)
p0

ggsave('F1b.new.png',p0,height=5,width=5)


```




# load Seurat data 
```{r}



# embedding of all fraction 

Seuratall=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat4/Seurat.all.rds')
emball=Seuratall@reductions$umap@cell.embeddings


f1=embeddingPlot(emball,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.2)+  
  theme(panel.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=0.3))
#  theme( panel.grid.major = element_blank(),
#   panel.grid.minor = element_blank(),
#    panel.border = element_blank(),
#   panel.background = element_blank()) 



ggsave('R1.Seurat.all2.embedding.png',f1,height=4,width=4)







# load Seurat data  Seurat embedding  
ST=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat2/Tumor.out.rds')

dim(ST@reductions$umap)
emb=ST@reductions$umap@cell.embeddings


f2=embeddingPlot(emb,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.2)+  
  theme(panel.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=0.3))


ggsave('R1.Seurat.Tumor.fraction.png',f2,height=4,width=4)




# Distal.fraction
ST=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat2/Noninvolved.out.rds')
DimPlot(ST, reduction = "umap", label = TRUE)

dim(ST@reductions$umap)
emb=ST@reductions$umap@cell.embeddings


f3=embeddingPlot(emb,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,min.cluster.size = 20,palette=pf,size=0.2)+  
  theme(panel.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=0.3))


ggsave('R1.Seurat.Distal.fraction.png',f3,height=4,width=4)


# Involved  fraction
ST=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat2/Involved.out.rds')
DimPlot(ST, reduction = "umap", label = TRUE)

dim(ST@reductions$umap)
emb=ST@reductions$umap@cell.embeddings


f4=embeddingPlot(emb,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.2,min.cluster.size = 20)+  
  theme(panel.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=0.3))


ggsave('R1.Seurat.Involved.fraction.png',f1,height=4,width=4)



# Benign fraction
ST=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat2/Whole.out.rds')
DimPlot(ST, reduction = "umap", label = TRUE)

dim(ST@reductions$umap)
emb=ST@reductions$umap@cell.embeddings


f5=embeddingPlot(emb,groups=typefc2,font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.2,min.cluster.size = 20)+  
  theme(panel.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, size=0.3))


ggsave('R1.Seurat.Whole.fraction.png',f1,height=4,width=4)



```


Seurat integration for all fraction 
```{r fig.width=5, fig.height=5}
f1

```

Seurat integration for individual fraction 

Bengin Distal
Involved Tumor
```{r fig.width=8, fig.height=8}


b=  cowplot::plot_grid(plotlist=list(f5,f3,f4,f2), ncol=2, nrow=2)

b
```

#   cell distance using Seurte result 
```{r fig.width=5, fig.height=5}

# load Seurte DE genes 
ST=readRDS('/home/meisl/Workplace/BMME/Revision/Seurat2/Tumor.out.rds')

allde=readRDS('Seurat.DE.rds')

# Seurte DE genes 
#allde=list()
#for (x in unique(cell)){
#  print(x)
#  allde[[x]]=FindMarkers(object = ST, ident.1 = x,group.by='cell')
#}

#saveRDS(allde,'Seurat.DE.rds')



gl1 <- unname(unlist(lapply(allde, function(x) {
  x=x[x[,'avg_logFC']>0,]
  x <- x[order(x$p_val),]
  head(rownames(x),n=120)
})))


gl1=unique(gl1)

# expression matrix of Seurat
t.exp=ST@assays$RNA@data
#dim(t.exp)

cname2=intersect(names(cellano),colnames(t.exp))
length(cname2)
t.exp=t(t.exp[gl1,cname2])
#dim(t.exp)

# Avearge expression per cell type 
cma=apply(as.matrix(t.exp),2,function(x) tapply(x,cellano[rownames(t.exp)],mean))
cma.cor.spearman <- cor(as.matrix(t(cma)),method='spearman')
hc.spearman <- hclust(as.dist(1-cma.cor.spearman))


pdf('Seurte expression distance.pdf',height=4,width=4.5)
plot(hc.spearman,main='Seurte expression distance',xlab='',ylab='',sub="")
dev.off()

plot(hc.spearman,main='Seurte expression distance',xlab='',ylab='',sub="")

```

#   cell distance using conos result 
```{r fig.width=5, fig.height=5}
##
#de_conos <- p2T$getDifferentialGenes(groups=cellano)


gl1 <- unname(unlist(lapply(de_conos, function(x) {
  x=x[x$Z>0,]
  x <- x[order(x$Z,decreasing=T),]
  head(rownames(x),n=120)
})))

#length(unique(gl1))
gl1=unique(gl1)


t.exp=p2T$counts[names(cellano),gl1]

cma2=apply(t.exp,2,function(x) tapply(x,cellano,mean))
#dim(cma2)

cma.cor.spearman <- cor(as.matrix(t(cma2)),method='spearman')
hc.spearman2 <- hclust(as.dist(1-cma.cor.spearman))
plot(hc.spearman2)


pdf('Conos expression distance.pdf',height=4,width=4.5)
plot(hc.spearman2,main='Conos expression distance',xlab='',ylab='',sub="")
dev.off()

plot(hc.spearman2,main='Conos expression distance',xlab='',ylab='',sub="")

```










# score projenitor downstream branches 
```{r fig.width=5, fig.height=5}

getScore_Cell=function(aexp,gs,group,magnitude.normal=NULL,rscore=NULL){
  # exp colname is gene 
  
  gs=intersect(gs,rownames(aexp))
  cname=names(group)
  exp=as.matrix(aexp[gs,cname])
  
  exp=exp[rowMeans(exp)!=0,] # remove unexpressed gene 
  exp=exp[,colSums(exp)!=0]
  cname=intersect(cname,colnames(exp))
  
  if (!is.null(magnitude.normal)){
    exp=apply(exp,2,function(x) x/max(x))
  }
  
  score=rowMeans(t(exp))
  return(score) 
}


sel=c("Erythroid","Immature B","Monocyte prog")

P2allde=p2all$getDifferentialGenes(groups=Toch(typefc2[typefc2 %in% sel]))
names(P2allde)

```

downstream branches singature gene 
```{r fig.width=5, fig.height=5}
f=DEheatmap(p2all,Toch(typefc2[typefc2 %in% sel[1:3]]),'HSC.Score')
plot(f$gtable)

```

# embedding of Progenitors and downstream branches
```{r fig.width=5, fig.height=5}
cname=names(typefc2[typefc2=='Progenitors'])
emb=t(scon$embedding[,cname])

# remove outliner 
#plot(emb)
emb=emb[emb[,2]>0,]
cname=rownames(emb)


cname_ano=names(typefc2[typefc2 %in% c(sel[1:3],'Progenitors')])
cname_ano=t(scon$embedding[,cname_ano]) %>% .[.[,2]>0,] %>% rownames()


p0 <- scon$plotGraph(raster=T,groups=typefc2[cname_ano],font.size=c(3.4,5),alpha=0.08,plot.na=F,palette=pf,size=0.8,title='cell')
p0
```


# Scoring projenitors using downstream branches signature genes 
```{r fig.width=8, fig.height=8}
aexp=t(p2all$counts)
allg_pfilter=lapply(P2allde,function(x) {
  x=x[x$Z>0,]
  #rownames(x)=x[,1]
  # x=x[order(x$Z,decreasing=TRUE),][1:200,]
  gs=rownames(x)[1:100]
  gs[!is.na(gs)]
})

lis=list()

lis[['cell']]=p0

for (ii in sel[1:3]){
  rm=setdiff(sel,ii)
  gs=setdiff(allg_p3[[ii]],do.call(c,allg_p3[rm]))
#  length(gs)
  gs=intersect(gs,allg_pfilter[[ii]])
  gs=gs[1:15]
  gs=gs[!is.na(gs)]
  
  gs=allg_pfilter[[ii]][1:20]
  finalg[[i]]=gs
 # print(gs)
  score=getScore_Cell(aexp,gs,Toch(typefc2)[cname])
  
 # plot(density(score))
  q1=quantile(score,0.99)
  score[score>q1]=q1
  
  col <- pagoda2:::val2col(score,gradient.range.quantile = 0.9,gradientPalette = colorRampPalette(c(rep('grey85',3),'red'))(1024))
  p3 <- scon$plotGraph(alpha=0.1,size=0.3,colors=col,plot.na=F,title=ii)

  ggsave(paste(ii,'score2.png',sep=''),p3,width = 3,height=3.1)
  lis[[ii]]=p3
  
  
}

lrow=2
lcol=2
b=  cowplot::plot_grid(plotlist=lis, ncol=lrow, nrow=lcol)

fout='Q1.HSC.png'
ggsave(fout,b,width = 2.3*lrow,height=2.3*lcol)


b

```


# downstream branches signature gene expression in projenitors
```{r fig.width=9, fig.height=12}


gs=c('CD79A','TCL1A','VPREB3','MPO','AZU1','LYZ','HBB','HBD','CA1','CD164','SPINK2','SOX4')

#
appname='HSC.marker.png'
lrow=3
lcol=4
alpha=0.5


lis=list()
for (gene in gs){
  #print(gene)
  t=p2all$counts[cname,gene]
  a=scon$plotGraph(colors =t,title=gene,alpha =0.8,size=0.01,plot.na=F)
  lis[[gene]]=a
}
b=  cowplot::plot_grid(plotlist=lis, ncol=lrow, nrow=lcol)
fout=paste(appname,'expConos.png',sep='')
ggsave(fout,b,width = 2.3*lrow,height=2.3*lcol)

b
```



